---
name: Philipp
surname: Bayer
position: "Research Officer, Computational Biology"
address: "OceanOmics, Minderoo Foundation"
email: "pbayer@minderoo.org"
github: philippbayer
linkedin: philippbayer
date: "`r format(Sys.time(), '%B %Y')`"
output: 
  vitae::awesomecv:
    page_total: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
library(vitae)
library(tidyverse)
library(RefManageR)
library(tibble)
```

# About me

I work at the intersection of genomics, data science, machine learning and AI, high-performance computing, teaching, and environmental science, drawing on more than a decade of experience at the cutting edge of genomics to push what is possible in environmental and DNA data analysis. I am member of two scientific advisory panels, have co-supervised more than twenty PhD and MSc students to completion, published roughly 130 research publications, co-edited a book, and am active in machine learning, eDNA, and open source communities.  
At Minderoo OceanOmics I am focusing on pushing the boundary of what is possible using environmental DNA (eDNA) data using genomics and artificial intelligence.


# Employment

```{r}
tribble(
  ~Year, ~Role, ~Location, ~Institution, ~Details,
  "2022 to current", "Research Officer, Computational Biology", "Perth", "Minderoo OceanOmics", "I lead research efforts using interpretable artificial intelligence and genomics to improve how we measure, understand, and protect the ocean.",
  "2021 to 2022", "DECRA Fellow", "UWA, Perth", "UWA", "I used interpretable AI to study mechanisms of gene loss and birth to learn how to avoid loss of agronomically important genes. I co-developed a new MSc in Bioinformatics at UWA and designed and taught two units.",
  "2017 to 2020", "Forrest Fellow", "UWA, Perth", "UWA", "I was one of three inaugural Forrest Fellows. I worked on genomics of complex plants with Forrest Foundation support. I taught one bioinformatics unit.",
  "2015 to 2017", "Postdoctoral researcher", "UWA, Perth", "UWA", "I researched genetics of complex plants with a focus on canola and wheat in the Edwards lab.",
) %>% 
    brief_entries(glue::glue("\\textbf{<Role>}, <Institution> \\newline <Details>", .open = "<", .close = ">"), Year, .protect = FALSE)
```



# Education

```{r}
tribble(
  ~ Degree, ~ Year, ~ Institution, ~loc, ~ details,
  "PhD", "2012 to 2016", "University of Queensland", "Brisbane, Australia", 'Applied bioinformatics PhD in the Edwards group. I developed computational pipeline SkimGBS for higher resolution genotyping while working extensively with industry on plant breeding pipelines (Bayer CropScience/BASF).', 
  "Master of IT", "2010 to 2012", "Bond University", "Gold Coast, Australia", "Focused on coding and business IT. 5x Top of class, 3x Vice-Chancellor List of Academic Excellence, 1x IT Award Academic Excellence. Graduated with High Distinction. John Oglethorpe Medal for highest GPA of all IT students graduating.",
  "Bachelor of Life Sciences", "2006 to 2009", "University of Muenster", "Muenster, Germany", 'Studied general life sciences with a focus on microbiology. In my final project I worked on EST-based differential gene expression in seagrasses.',
) %>% 
  brief_entries(glue::glue("\\textbf{<Degree>}, <Institution>, <loc> \\newline <details>", .open = "<", .close = ">"), Year, .protect = FALSE)
```

\newpage

# Recent roles

```{r}
tribble(
  ~Year, ~Role, ~Organisation, ~Details,
  "2023", "Founder, WA eDNA meetup", "Minderoo Foundation", "Meetup to connect eDNA practitioners in Western Australia",
  "2022", "Member, Scientific Advisory Panel", "UWA Data Institute", "Oversee and advise on ongoing industry data collaborations across WA.",
  "2022", "Founding member", "UWA Centre for Applied Bioinformatics", "Co-founder and lead of teaching.",
  "2021", "Member, Scientific Advisory Panel Machine Learning", "ARDC", "Advise on ongoing machine learning projects supported by the ARDC.",
  "2021", "Member, GRDC Machine Learning Technical Consultation Group (ML TCG),", "GRDC", "Member of the ML technical consultation group for GRDC-funded ML projects.",
  "2017", "Hacky Hour Founder", "UWA, Perth", "Founded a weekly get-together of researchers and staff working with programming and data, doubles as a help-desk for students with programming problems.",
  "2013-2023", "Certified Carpentries Instructor", "Australia", "Certified Software Carpentry and Data Carpentry instructor",
) %>% 
  brief_entries(glue::glue("\\textbf{<Role>} \\newline <Details>", .open = "<", .close = ">"), Year, .protect = FALSE)
```


# Recent presentations

I have given a total of 40 public presentations.

```{r}
tribble(
  ~Year, ~Role, ~Organisation, ~Details,
  "2024", "What we learned using LLMs in environmental DNA", 'ACS WA conference', 'Invited talk',
  "2024", 'How we can use social media to protect the oceans', 'TEDx Kings Park', 'TEDx!',
  "2024", 'Towards better standards for eDNA taxonomic classification', 'Curtin University', 'Invited talk',
  "2023", 'My Journey to a Career in Machine Learning; All My Mistakes You Will Hopefully Avoid', "WADSIH Data and AI for Business Conference", 'Invited talk',
  "2023", 'Unlocking the power of large language models', 'Perth Machine Learning Group', 'Invited talk',
  "2023", 'My journey towards data skills', 'WA Data Science Innovation Hub (WADSIH)', 'Invited talk',  
  "2022", 'Ask us anything - Advice on how to use machine learning to tackle your project', 'ResBaz/Perth Machine Learning Group', 'Panel Q&A',
  '2022', 'Career Pathways Event - how I got here', 'Telethon Kids Institute', 'Invited talk',
  "2022", 'Bioinformatics showcase', 'ResBaz', 'Live demo of bioinformatics analysis',
  "2021", "Machine learning in bioinformatics – where are we and whatʼs next?", "CCDM/Curtin University", "Invited talk",
  "2021", "Machine learning in plant breeding and bioinformatics", "Cinvestav/online", "Invited talk",
  "2021", "Future-ready crops for a changing climate: the role of bioinformatics", "UWA DVCR Forrest Fellow series", "Invited talk",
  "2021", "Bioinformatics at scale Q and A", "Pawsey Supercomputing Centre", "Invited",
  "2021", "Interpretable Machine Learning in Bioinformatics", "ABACBS online seminars", "Invited talk",
) %>% 
  brief_entries(glue::glue("\\textbf{<Organisation>} - <Role>", .open = "<", .close = ">"), Year, .protect = FALSE)
```


# Academic funding

I have received a total of $1.5 million in academic research funding.

```{r}
tribble(
  ~Year, ~Role, ~Organisation, ~Details,
  "2020", "Grant: ARC Discovery Early Career Research Award", "ARC", "Awarded DECRA for 2021-2023. Total funding \\$866k which includes a \\$419k contribution from UWA.",
  "2020", "Grant: Identifying genetic contributors to canola blackleg resistance in the presence of environmental effects using Machine Learnin", "GRDC", "With Prof. Dave Edwards, Prof. Mohammed Bennamoun, Prof. Farid Boussaid, Prof. Jacqueline Batley. Total funding: AUSD 309,524.",
  "2020", "Grant: Machine Learning - Project E: Deep Learning for early detection and
classification of crop disease and stress", "GRDC", "With Prof. Mohammed Bennamoun, Prof. Farid Boussaid, Prof. Dave Edwards, Dr. Nic Taylor. Total funding: AUSD 344,971."
) %>% 
  brief_entries(glue::glue("\\textbf{<Role>} \\newline <Details>", .open = "<", .close = ">"), paste0(Year, "\\newline"), .protect = FALSE)
```

\newpage

# Selected publications

These represent the most impactful publications out of roughly 130 publications.

```{r}
d <- as_tibble(ReadBib('CV/important_papers.bib'))
d %>% 
  mutate(author = as.character(map(strsplit(author, split = " and "), 1))) %>% # pull out first author only
  mutate(author = gsub("^.* ", "", author)) %>% # get author last name
  arrange(desc(year), author) %>% rowid_to_column() %>% 
  brief_entries(glue::glue("\\textbf{<author>} \\textit{et al.} (<year>). <title>. \\textit{<journal>}, \\textit{<volume>}, <number>:<pages>.", .open = "<", .close = ">"), when=rowid, .protect = FALSE)
```

