---
name: Philipp
surname: Bayer
position: "DECRA Fellow"
address: "School of Biological Sciences, University of Western Australia"
www: philipp-bayer.net
email: "philippbay@gmail.com"
twitter: philippbayer
github: philippbayer
linkedin: philippbayer
date: "`r format(Sys.time(), '%B %Y')`"
output: 
  vitae::awesomecv:
    page_total: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
library(vitae)
library(tidyverse)
```

# About me

I use computers and large data sets to answer biological questions, especially in crops. I use genomics, pangenomics, k-mer/SNP/QTL-association studies, biostatistics, and interpretable machine learning to find new approaches to breed climate change-ready crops.

# Education

```{r}
library(tibble)
tribble(
  ~ Degree, ~ Year, ~ Institution, ~ Where,
  "PhD", "2012-2016", "University of Queensland", "Brisbane, Australia",
  "Master of IT", "2010-2012", "Bond University", "Gold Coast, Australia",
  "Bachelor of Life Sciences", "2006-2009", "University of Muenster", "Muenster, Germany"
) %>% 
  detailed_entries(Degree, Year, Institution, Where)
```

# Employment

```{r}
tribble(
  ~Year, ~Role, ~Location, ~Details,
  "2021-2023", "DECRA Fellow", "Perth, UWA", "My first step towards an independent lab with funding for the first PhD-student primarily supervised by me and with AUSD 448,781 in government and AUSD 418,772 UWA funding. I am modeling mechanisms of gene loss and birth in crops to learn where new genes come from, and how to avoid loss of agronomically important genes.",
  "2017-2020", "Forrest Fellow", "Perth, UWA", "One of three inaugural Forrest Fellows. Continued work on genomics of complex plants with Forrest Foundation support. Supervised three PhD students and four MSc students to completion.",
  "2015-2017", "Postdoctoral researcher", "Perth, UWA", "Edwards Lab. Researched genetics of complex plants with a focus on canola and wheat. Worked closely with industry partners to improve their breeding programs. Supervised two interns, co-supervised four PhD students and one MSc student, system administrator for the local computational infrastructure and group data manager.",
) %>% 
  detailed_entries(when=Year, with=Role,  where=Location, what=Details)
```


# Experience

```{r}
tribble(
  ~Year, ~Role, ~Organisation, ~Details,
  "2021", "Host, ABACBS conference Perth Hub", "Perth", "Hosted the Perth hub of the yearly ABACBS conference, chaired genomics session",
  "2021-Current", "Member, Scientific Advisory Panel Machine Learning", "ARDC", "Member of the scientific advisory panel for ongoing machine learning projects supported by the ARDC.",
  "2012-2018", "Research Collaboration with Bayer CropScience, later BASF", "Ghent, Belgium", "Continued collaboration with Bayer CropScience on their plant breeding projects.",
  "2011-Current", "Co-founder openSNP.org", "Germany/Australia", "A project for customers of genotyping companies like 23andMe to share their data with scientists around the world, for free. Partially wrote and  maintain the siteʼs uby on Rails code-base, interact and manage with the community of 5000 users, administration of the siteʼs servers, and supervision of contributors",
  "2013-Current", "Certified Carpentries Instructor", "Australia", "Certified Software Carpentry and Data Carpentry instructor. The Carpentries is a non-profit organization teaching programming to scientists around the world.",
   "2018", "Research Bazaar Organising Committee", "UWA, Perth",  "ResBaz is a world-wide three-day festival promoting digital literacy. As member of the organising committee I searched for helpers and teachers, drafted the timeplan,designed the web page, raised funding, succeeded in getting government MP to hold the keynote",
    "2017-Current", "Hacky Hour Funder", "UWA, Perth", "Founded the Hacky Hour at UWA, a weekly get-together of researchers and staff working with programming and data, doubles as a help-desk for students with programming problems.",
    "2017-2019", "Mozilla Open Science Leadership Mentor", "Online", "Mentored open source programmers and researchers on how to streamline and grow open source and open science projects under the umbrella of Mozilla",
    "2016-2019", "EMBLR-ABR Head of Nodes member, Open Science Special Interest Group member", "UWA, Perth", "EMBL-ABR was an Australian-wide network supporting the technical needs of life sciences researchers.",
     "2016-2017", "COMBINE WA Representative", "UWA, Perth", "COMBINE is the student and early career researcher subcommittee of the Australian Bioinformatics and Computational Biology Society (ABACBS). As the local representative I organise or help organise workshops and regular networking events."
) %>% 
  detailed_entries(when=Year, with=Role, where=Organisation, what=Details)
```

# Awards and Funding

```{r}
tribble(
  ~Year, ~Role, ~Organisation, ~Details,
  "2021-2023", "Grant: ARC Discovery Early Career Research Award", "ARC", "Awarded DECRA for 2021-2023. Total funding: AUSD 448,781 and ASUD 418,772 in UWA funding.",
  "2020-2022", "Grant: Identifying genetic contributors to canola blackleg resistance in the
presence of environmental effects using Machine Learnin", "GRDC", "With Prof. Dave Edwards, Prof. Mohammed Bennamoun, Prof. Farid Boussaid, Prof. Jacqueline Batley. Total funding: AUSD 309,524.",
  "2020-2022", "Grant: Machine Learning - Project E: Deep Learning for early detection and
classification of crop disease and stress", "GRDC", "With Prof. Mohammed Bennamoun, Prof. Farid Boussaid, Prof. Dave Edwards, Dr. Nic Taylor. Total funding: AUSD 344,971.",
  "2019", "Woodside Early Career Scientist of the Year, finalist", "WA", "Finalist in Premierʼs Science Awards 2019",
  "2018", "Rising Stars nomination", "UWA, Perth", "Two early career researchers per UWA research school were nominated for Rising Stars, a university-wide event where researchers introduce a public audience to their research.",
   "2018", "Forrest Research Foundation Non-stipendiary Fellowship", "UWA, Perth",  "Three year fellowship to pursue research at UWA, part of the three inaugural Forrest Fellows",
    "2017", "UWA Research Collaboration Award ", "UWA, Perth", "AUSD 28,100 to fund a seagrass microbiome sequencing project.",
    "2014", "GRDC Travel Award", "GRDC", "Travel Cost Scholarship to travel to PAG, USA.",
    "2014", "UQ SAFS Travel Award", "UQ, Brisbane", "Travel Cost Scholarship to travel to PAG, USA.",
    "2013", "Bayer Grants4Apps", "Bayer", "Grant to cover openSNP running costs",
    "2011-2014", "Two postgraduate scholarships", "UQ, Brisbane", "My PhD was supported by two scholarships from UQ for tuition fees and living costs.",
    "2012", "First place in PLOS/Mendeley Binary Challenge", "openSNP.org", "Won first price in a competition aimed towards the advancement of open science",
    "2009-2001", "Master IT", "Bond University", "5x Top of class, 3x Vice-Chancellorʼs List for Academic Excellence, 1x IT Award Academic Excellence. Graduated with High Distinction. John Oglethorpe Medal for highest GPA of all IT students graduating.",
) %>% 
  detailed_entries(when=Year, with=Role, where=Organisation, what=Details)
```

# Teaching

```{r}
tribble(
  ~Year, ~Role, ~Organisation, ~Details,
  "2017-Current", "University teaching", "UWA, Perth", "Organised new MSc Bioinformatics with new unit, SCIE5003 (advanced bioinformatics). Developed content of SCIE5003 and SCIE4002, taught into both units.",
  "2021", "Introduction to tidyverse and purrr", "ASI R Workshop", "Introducing RStudio, project organisation, basic tidyverse, loops, functions, maps",
  "2019", "Introduction to genomics on the command line", "ResBaz, Curtin", "Introduction to the command line, bioinformatics analyses and pipelines, and basic SNP analysis in R.",
  "2019", "Introduction to tidyverse and caret in R", "School of Human Sciences, UWA", "Introduction to R, tidyverse, ggplot2, caret, and basic statistics in R. Taught over two
days.",
  "2018", "Introduction to modern R", "TKI, Perth", "Introduction to R, tidyverse, ggplot2, and basic statistics approaches in R. Taught over two days.",
   "2018", "Data Carpentry workshop", "ResBaz, UWA",  "Introduction to genomics and shell. Part of the planning committee.",
    "2017", "Software Carpentry workshop", "ResBaz, Curtin", "Introduction to data manipulation using Python.",
    "2016", "Data Carpentry workshop and host", "UWA, Perth", "Hosted, planned, and set up the first Data Carpentry workshop at UWA, taught best data management practices.",
    "2016", "Software Carpentry workshop", "ResBaz, Murdoch Uni", "Introduction to Python and git",
    "2016", "Software Carpentry workshop", "Curtin Uni, Perth", "Introduction to Python and git",
    "2016", "Software Carpentry workshop and host", "UQ, Brisbane", "Hosted, planned, and taught into the first Sofware Carpentry workshop at UQ.",
    "2014", "Software Carpentry workshop", "Sydney", "Basic to intermediate Python.",
    "2014", "Software Carpentry workshop", "PyCon AU/UQ, Brisbane", "Basic to intermediate Python",
    "2013", "Software Carpentry workshop", "Adelaide", "Basic to intermediate Python as assistant",
    "2009-2011", "Programming tutor", "Bond University", "Tutored students in Intro to Programming (Java), Database Management (Oracle/MySQL) and Networks & Applications, held several all-day refresher courses before exams",
) %>% 
  detailed_entries(when=Year, with=Role, where=Organisation, what=Details)
```


# Presentations
```{r}
tribble(
  ~Year, ~Role, ~Organisation, ~Details,
  "2021", "Machine learning in bioinformatics – where are we and whatʼs next?", "CCDM/Curtin University", "Invited talk",
  "2021", "Machine learning in plant breeding and bioinformatics", "Cinvestav/online", "Invited talk",
  "2021", "Future-ready crops for a changing climate: the role of bioinformatics", "UWA DVCR Forrest Fellow series", "Invited talk",
  "2021", "Bioinformatics at scale Q & A", "Pawsey Supercomputing Centre", "Invited",
  "2021", "Interpretable Machine Learning in Bioinformatics", "ABACBS online seminars", "Invited talk",
  "2020", "Our machine learning technical stack", "GRDC Tech Seminars", "",
  "2020", "Predicting Gene Loss in Plants: Lessons Learned from Laptop-Scale Data", "PAG Conference, San Diego", "",
  "2019", "Eukaryotic pangenomics: where we've been, where we're going", "Bayliss Seminar Series", "Invited",
  "2019", "Assembling complex plant genomes – things I wish someone would have told
me earlier", "AGRF Seminar Series", "Invited",
  "2019", "Helping Biologists Make Sense of Plant Variant and Annotation Data", "PAG Conference, San Diego", "",
  "2018", "Feeding the future world: safe-guarding Australiaʼs food bowl in a changing
climate", "Rising Stars UWA", "",
  "2018", "From QTLs to candidate genes, or: There and Back Again", "Institute of Ag seminars", "Invited",
  "2018", "The path of least resistance (genes) - mining plant genomes for disease resis-
tance", "COMBINE/Pawsey symposium", "Invited",
  "2018", "Early Career Researcher Panel - What have I learnt at the beginning of my re-
search career?", "Combined Biological Sciences Meeting", "",
  "2018", "ScienceCafe - STEM outreach aimed at year 10 students", "UWA", "",
  "2017", "The future of wheat research", "UWA" ,"", 
  "2017", "Skipping the assembly step – what we can learn from looking at sequences
directly", "Pawsey Roadshow", "Invited",
  "2017", "The State of Bioinformatics in High Performance Computing in 2017", "HPCAC Conference, Perth", "",
  "2017", "Towards better plant breeding at UWA", "COMBINE, Perth","",
  "2017", "Improving Plant Breeding using KNetMiner", "PAG, San Diego", "",
  "2016", "Towards a canola pan-genome: cautionary tales from the assembly bench", "CCDM, Curtin", "Invited talk",
  "2016", "Sharing Experience: What Can We Learn from Each Other Developing Plant
Informatics Systems", "PAG, San Diego", "",
  "2015", "Assessing and validating the amphidiploid genome of Brassica napus using
genotyping by sequencing", "PAG, San Diego", "",
  "2015", "Using skim-based genotyping by sequencing for trait association and QTL
cloning in Brassica napus", "PAG, San Diego", "",
  "2014", "Assembling and validating the genome of the Brassica napus using skim-
based genotyping by sequencing", "UQ, GenGen Seminars", "",
  "2012", "openSNP: Crowdsourcing Genome Wide Association Studies", "Chaos Communication Congress, Berlin", "",
) %>% 
  detailed_entries(when=Year, with=Role, where=Organisation, what=Details)
```

# Publications

```{r}
scholar::get_publications("GOL15xQAAAAJ") %>% 
    dplyr::filter(!is.na(year)) %>%  # filter out weird trash
    filter(!grepl("Plant and Animal Genome", journal)) %>%  # filter out talks
    arrange(desc(year)) %>% # newest papers first
    detailed_entries(
        with=title,
        when=year,
        what=author,
        where=journal
        )
```

