---
name: Philipp
surname: Bayer
position: "Research Officer, AI"
address: "OceanOmics, Minderoo Foundation"
email: "pbayer@minderoo.org"
twitter: philippbayer
github: philippbayer
linkedin: philippbayer
date: "`r format(Sys.time(), '%B %Y')`"
output: 
  vitae::awesomecv:
    page_total: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
library(vitae)
library(tidyverse)
library(RefManageR)
```

# About me

I work at the intersection of genomics, data science, machine learning and AI, high-performance computing, teaching, and environmental science, drawing on more than a decade of experience at the cutting edge of genomics to push what is possible in environmental and DNA data analysis. I have co-supervised more than twenty PhD and MSc students to completion and am now focusing on building a team of AI experts in eDNA. 

# Education

```{r}
library(tibble)
tribble(
  ~ Degree, ~ Year, ~ Institution, ~loc, ~ details,
  "PhD", "2012 to 2016", "University of Queensland", "Brisbane, Australia", 'Applied bioinformatics PhD in the Edwards group. Developed computational pipeline SkimGBS for higher resolution genotyping. Worked extensively with industry (Bayer CropScience/BASF).', 
  "Master of IT", "2010 to 2012", "Bond University", "Gold Coast, Australia", "Focused on coding and business IT. 5x Top of class, 3x Vice-Chancellor List of Academic Excellence, 1x IT Award Academic Excellence. Graduated with High Distinction. John Oglethorpe Medal for highest GPA of all IT students graduating.",
  "Bachelor of Life Sciences", "2006 to 2009", "University of Muenster", "Muenster, Germany", 'Studied general life sciences with a focus on microbiology. In my final project I worked on EST-based differential gene expression in seagrasses.',
) %>% 
  brief_entries(glue::glue("\\textbf{<Degree>}, <Institution>, <loc> \\newline <details>", .open = "<", .close = ">"), Year, .protect = FALSE)
```

# Employment

```{r}
tribble(
  ~Year, ~Role, ~Location, ~Institution, ~Details,
  "2022 to current", "Research Officer, AI", "OceanOmics", "Minderoo", "I lead interpretable AI efforts using genomics and flow cytometry data from WA to assess ocean health.",
  "2021 to 2022", "DECRA Fellow", "UWA, Perth", "UWA", "I used interpretable AI to study mechanisms of gene loss and birth in crops to learn where new genes come from, and how to avoid loss of agronomically important genes. Co-developed a new MSc in Bioinformatics, designed and taught two units.",
  "2017 to 2020", "Forrest Fellow", "UWA, Perth", "UWA", "One of three inaugural Forrest Fellows. Worked on genomics of complex plants with Forrest Foundation support. Taught one bioinformatics unit.",
  "2015 to 2017", "Postdoctoral researcher", "UWA, Perth", "UWA", "Researched genetics of complex plants with a focus on canola and wheat in Edwards lab.",
) %>% 
    brief_entries(glue::glue("\\textbf{<Role>} \\newline <Details>", .open = "<", .close = ">"), Year, .protect = FALSE)
```


# Recent roles

```{r}
tribble(
  ~Year, ~Role, ~Organisation, ~Details,
  "2023", "Founder, WA eDNA meetup", "Minderoo Foundation", "Irregular meetup to connect eDNA practitioners in Western Australia",
  "2022", "Member, Scientific Advisory Panel", "UWA Data Institute", "Oversee and advise on ongoing industry data collaborations across WA.",
  "2022", "Founding member", "UWA Centre for Applied Bioinformatics", "Co-founder and lead of teaching.",
  "2021", "Member, Scientific Advisory Panel Machine Learning", "ARDC", "Advise on ongoing machine learning projects supported by the ARDC.",
    "2021", "Member, GRDC Machine Learning Technical Consultation Group (ML TCG),", "GRDC", "Member of the ML technical consultation group for GRDC-funded ML projects.",
  "2017", "Hacky Hour Founder", "UWA, Perth", "Founded a weekly get-together of researchers and staff working with programming and data, doubles as a help-desk for students with programming problems.",
  "2013", "Certified Carpentries Instructor", "Australia", "Certified Software Carpentry and Data Carpentry instructor",
) %>% 
  brief_entries(glue::glue("\\textbf{<Role>} \\newline <Details>", .open = "<", .close = ">"), Year, .protect = FALSE)
```

\newpage

# Recent presentations
```{r}
tribble(
  ~Year, ~Role, ~Organisation, ~Details,
  "2023", 'Unlocking the power of large language models', 'Perth Machine Learning Group', 'Invited talk',
  "2023", 'My journey towards data skills', 'WA Data Science Innovation Hub (WADSIH)', 'Invited talk',  
  "2022", 'Ask us anything - Advice on how to use machine learning to tackle your project', 'ResBaz/Perth Machine Learning Group', 'Panel Q&A',
  '2022', 'Career Pathways Event - how I got here', 'Telethon Kids Institute', 'Invited talk',
    "2022", 'Bioinformatics showcase', 'ResBaz', 'Live demo of bioinformatics analysis',
  "2021", "Machine learning in bioinformatics – where are we and whatʼs next?", "CCDM/Curtin University", "Invited talk",
  "2021", "Machine learning in plant breeding and bioinformatics", "Cinvestav/online", "Invited talk",
  "2021", "Future-ready crops for a changing climate: the role of bioinformatics", "UWA DVCR Forrest Fellow series", "Invited talk",
  "2021", "Bioinformatics at scale Q and A", "Pawsey Supercomputing Centre", "Invited",
  "2021", "Interpretable Machine Learning in Bioinformatics", "ABACBS online seminars", "Invited talk",
  "2020", "Our machine learning technical stack", "GRDC Tech Seminars", "",
  "2020", "Predicting Gene Loss in Plants: Lessons Learned from Laptop-Scale Data", "PAG Conference, San Diego", "",
  "2019", "Eukaryotic pangenomics: where we've been, where we're going", "Bayliss Seminar Series", "Invited",
  "2019", "Assembling complex plant genomes – things I wish someone would have told
me earlier", "AGRF Seminar Series", "Invited",
  "2019", "Helping Biologists Make Sense of Plant Variant and Annotation Data", "PAG Conference, San Diego", "",
) %>% 
  brief_entries(glue::glue("\\textbf{<Organisation>} - <Role>", .open = "<", .close = ">"), Year, .protect = FALSE)
```


# Academic funding

```{r}
tribble(
  ~Year, ~Role, ~Organisation, ~Details,
  "2020", "Grant: ARC Discovery Early Career Research Award", "ARC", "Awarded DECRA for 2021-2023. Total funding \\$866k which includes a \\$419k contribution from UWA.",
  "2020", "Grant: Identifying genetic contributors to canola blackleg resistance in the
presence of environmental effects using Machine Learnin", "GRDC", "With Prof. Dave Edwards, Prof. Mohammed Bennamoun, Prof. Farid Boussaid, Prof. Jacqueline Batley. Total funding: AUSD 309,524.",
  "2020", "Grant: Machine Learning - Project E: Deep Learning for early detection and
classification of crop disease and stress", "GRDC", "With Prof. Mohammed Bennamoun, Prof. Farid Boussaid, Prof. Dave Edwards, Dr. Nic Taylor. Total funding: AUSD 344,971.",
) %>% 
  brief_entries(glue::glue("\\textbf{<Role>} \\newline <Details>", .open = "<", .close = ">"), paste0(Year, "\\newline"), .protect = FALSE)
```

# Relevant publications

```{r}
d <- as_tibble(ReadBib('CV/important_papers.bib'))
d %>% 
  mutate(author = as.character(map(strsplit(author, split = " and "), 1))) %>% # pull out first author only
  mutate(author = gsub("^.* ", "", author)) %>% # get author last name
  arrange(desc(year), author) %>% rowid_to_column() %>% 
  brief_entries(glue::glue("\\textbf{<author>} \\textit{et al.} (<year>). <title>. \\textit{<journal>}, \\textit{<volume>}, <number>:<pages>.", .open = "<", .close = ">"), when=rowid, .protect = FALSE)
```

